A New Method of Modelling Tuneable Lasers with Functional Composition

### 2019-07-12 Thesis/Model/Model.tex
Finished corrections

\chapter{A New Model}
\label{chap:linear}
In this chapter we shall derive our new model and solve it analytically in the linear case. To accomplish this we shall use the ideas of the previous functional models \cite{cutler, siegman, kuizenga1970a, kuizenga1970b, kuizenga1970, martinez1984, martinez1985, burgoyne2014}. To alleviate some of the pitfalls mentioned in the previous chapter, we shall include all five processes involved in the modification of the pulse within the laser cavity (gain, nonlinearity, loss, dispersion, and modulation). In addition to this, the functional operations associated with each component will be derived from \eqref{eq:nlse},
\begin{align*}
\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + \frac{\beta_3}{6}\pdiff[3]{A}{T} + i \gamma |A|^2 A + \frac{1}{2}g(A) A - \alpha A,
\end{align*}
with the exception of the modulation in which we consider the exact functional form to be determined by the laser operator. \\

\section{Components}
We shall begin our analysis with the derivation of the functional operators for the five components. \\

\subsection{Gain}
\label{chap:gain}
Within the Er-doped gain fibre, the gain term is dominant, and equation~\eqref{eq:nlse} reduces to
\begin{align}
\label{eq:gainde}
\pdiff{A}{z} &= \frac{1}{2} g(A),
\end{align}
where $g(A)$ takes the form \cite{bohun, burgoyne2014, hausbook, haus1975, haus1991, haus1992, haus2000, kartner, peng, shtyrina, silfvast, usechak, yarutkina}
\begin{align}
\label{eq:energy}
	g(A) &= \frac{g_0}{1 + E / \Es} A,& E &= \int_{-\infty}^\infty |A|^2 \, \df T,
\end{align}
where $g_0$ is a small signal gain, $E$ is the energy of the pulse, and $\Es$ is the energy at which the gain begins to saturate.
Multiplying~\eqref{eq:gainde} by $\bar{A}$, the complex conjugate of $A$, yields
\begin{align*}
	\bar{A} \pdiff{A}{z} = \frac{1}{2}\frac{g_0 |A|^2}{1 + E / \Es},
\end{align*}
adding this to its complex conjugate and integrating over $T$ gives
%\begin{align*}
%	\diff{|A|^2}{z} &= \frac{g_0 |A|^2}{1 + E / \Es}.
%\end{align*}
%After integrating this becomes
\begin{align}
\label{diffez}
	\diff{E}{z} &= \frac{g_0 E}{1 + E / \Es}.
\end{align}
For $E \ll \Es$ the energy will grow exponentially, whereas for $E \gg \Es$ the gain has saturated and so the growth is linear. To obtain a closed form solution, \eqref{diffez} is integrated over a gain fibre of length $z$ and assuming the energy increases from $E$ to $E_{\textrm{out}}$, then
\begin{align*}
	g_0 z = \log\frac{E_{\textrm{out}}}{E} + \frac{E_{\textrm{out}}-E}{\Es},
\end{align*}
and by exponentiating, rearranging, and applying $W$, the Lambert $W$ function\footnote{See Appendix \ref{chap:lambertw}.},
\begin{align*}
W \left(\frac{E}{\Es} \textrm{e}^{E/\Es} \textrm{e}^{g_0 z}\right) &= W\left(\frac{E_{\textrm{out}}}{\Es} \textrm{e}^{E_{\textrm{out}}/\Es}\right) = \frac{E_{\textrm{out}}}{\Es},
\end{align*}
by \eqref{eq:lambertw}. This results in the closed form expression
\begin{align*}
E_{\textrm{out}}(z) = \Es W \left( \frac{E}{\Es} \textrm{e}^{E/\Es} \textrm{e}^{g_0 z} \right)
\end{align*}
with the desired property that $E_{\textrm{out}}(0)=E$. Furthermore, since $E \sim |A|^2$, the gain in terms of the amplitude is given by
\begin{align*}
G(A;E) &= \left(\frac{E_{\textrm{out}}(L_g)}{E}\right)^{1/2}A = \left( \frac{\Es}{E} W \left( \frac{E}{\Es} \textrm{e}^{E/\Es} \textrm{e}^{g_0 L_g} \right) \right)^{1/2} A,
\end{align*}
where $L_g$ is the length of the gain fibre.

\subsection{Fibre Nonlinearity}
The nonlinearity of the fibre arises from the parameter $\gamma$; in regions where this effect is dominant expression~\eqref{eq:nlse} becomes
\begin{align}
\label{eq:fibrediff}
	\pdiff{A}{z} - i \gamma |A|^2 A = 0.
\end{align}
This expression can be manipulated in a similar manner to the gain to show $\frac{\partial}{\partial z} |A|^2 = 0$, suggesting that $A(z,T) = A_0(T) e^{i \varphi(T,z)}$. Substituting this representation into~\eqref{eq:fibrediff} and setting $\varphi(T,0)=0$ gives $\varphi(T,z) = \gamma |A|^2 z$. For a fibre of length $L_f$ the effect of the nonlinearity is thus
\begin{align*}
F(A) &= A \textrm{e}^{i \gamma |A|^2 L_f},
\end{align*}
this is also frequently called the Kerr nonlinearity or Kerr effect after John Kerr who discovered the effect in 1875. As we shall see in Section \ref{chap:energy}, this is what is responsible for self-phase modulation and ultimately the degradation of the pulse \cite{desurvire, dunlop, martinez1984, tamura}.

\subsection{Loss}
Two sources of loss exist within the laser circuit: the loss due to the output coupler and the optical loss due to absorption and scattering. The first case is simply a scalar multiplication depending on the reflectivity of the output coupler. The loss due to absorption and scattering can be derived from \eqref{eq:nlse},
\begin{align*}
\pdiff{A}{z} = - \alpha A.
\end{align*}
Combining these two effects give a loss that takes the form
\begin{align*}
L(A) &= (1 - R) \textrm{e}^{- \alpha L}A,
\end{align*}
where $R$ is the reflectivity of the output coupler, and $L$ is the total length of the laser circuit.

\subsection{Dispersion}
Within the laser cavity, the dispersion is dominated by the CFBG. In comparison, the dispersion due to the fibre is negligible. The dispersive terms of \eqref{eq:nlse} give
\begin{align}
\label{eq:dispde}
	\pdiff{A}{z} = -i \frac{\beta_2}{2} \pdiff[2]{A}{T} + \frac{\beta_3}{6} \pdiff[3]{A}{T},
\end{align}
and since dispersion acts in the frequency domain it is convenient to take the Fourier transform of~\eqref{eq:dispde} \cite{debnath, gradshteyn}, giving
\begin{align*}
	\pdiff{}{z}\FT{A} &= i\frac{\omega^2}{2}\left(\beta_2 - \frac{\beta_3}{3} \omega\right) \FT{A}.
\end{align*}
The effect of dispersion is then
\begin{align*}
D(A) &= \FTi{\textrm{e}^{i \omega^2 L_D(\beta_2 - \beta_3 \omega/3)/2} \FT{A}}.
\end{align*}
For a highly dispersive media the third order effects may need to be considered~\cite{agrawal2013, litchinitser}. However, for simplicity and because of the nature of the grating, the third order effects will be neglected \cite{agrawal2013, ferreira}---we set $\beta_3=0$ for the subsequent analysis. \\

\subsection{Modulation}
In the master equation of mode-locking, the amount of modulation is characterized by the parameter $M_s$ through the term $\frac{1}{2} M_s T^2 A$. In this new model, the modulation is considered to be applied externally through its action on the spectrum and for simplicity the representation is taken as the Gaussian
\begin{align*}
M(A) &= \textrm{e}^{-T^2 / 2 T_M^2} A,
\end{align*}
where $T_M$ is a characteristic width of the modulation. \\

Despite assuming a Gaussian modulation the following solution can be generalized to any modulation function in $L^2(\mathbb{R})$ since Gaussians in fact span $L^2(\mathbb{R})$ as will be shown shortly. \\

\input{./Appendices/Hermite}

@@

The structure of each process of the laser can be better understood by re-scaling the time, energy, and amplitude. Specifically, the time shall be scaled by the characteristic modulation time which is related to the pulse duration, the energy by the saturation energy, and the amplitude will be scaled so that it is consistent with the energy:
\begin{align*}
	T &= T_M \widetilde{T},& E &= \Es \widetilde{E},& A &= \left( \frac{\Es}{T_M} \right)^{1/2} \widetilde{A}.
\end{align*}
The new process maps, after dropping the tildes, become
\begin{subequations}
\label{eq:effects}
\begin{align}
\label{eq:gain}
G(A) &= \left(E^{-1} W \left( a E \textrm{e}^{E}\right) \right)^{1/2} A, \\
\label{eq:fibre}
F(A) &= A \textrm{e}^{i b |A|^2}, \\
\label{eq:loss}
L(A) &= h A, \\
\label{eq:disp}
D(A) &= \FTi{\textrm{e}^{i s^2 \omega^2} \FT{A}}, \\
\label{eq:mod}
M(A) &= \textrm{e}^{-T^2 / 2} A,
\end{align}
\end{subequations}
with the four dimensionless parameters (see Table~\ref{tab:values})
\begin{align*}
	a &= \textrm{e}^{g_0 L_g} \sim 8 \times 10^3,& s &= \sqrt{\frac{\beta_2 L_D}{2 T_M^2}} \sim 0.2, \\
	b &= \gamma L_f \frac{\Es}{T_M} \sim 1,& h &= (1 - R) \textrm{e}^{-\alpha L} \sim 0.04,
\end{align*}
which characterize the behaviour of the laser.

\begin{table}[tbp]
\centering
\begin{tabular}{lcll}
Parameter & Symbol & Value & Sources \\
\noalign{\global\arrayrulewidth=1.5pt}\hline
Absorption of Fibre & $\alpha$ & $0.01$--$0.3\text{ m}^{-1}$  & \cite{burgoyneemail, shtyrina, usechak, yarutkina} \\
Fibre Dispersion & $\beta_2^f$ & $-50$--$50 \text{ ps}^2/ \text{km}$ & \cite{agrawal2002, agrawal2013, burgoyne2014, litchinitser, peng, yarutkina} \\
Fibre Nonlinearity & $\gamma$ & $0.001$--$0.01 \text{ W}^{-1} \text{m}^{-1}$ & \cite{agrawal2013, usechak, yarutkina} \\
Grating Dispersion & $\beta_2^g L_D$ & $10$--$2000 \text{ ps}^2$ & \cite{agrawal2002, agrawal2013, burgoyne2014, li} \\
Length of Cavity & $L$ & $10$--$100 \text{ m}$ & \cite{burgoyneemail, peng} \\
Length of Fibre & $L_f$ & $0.15$--$1 \text{ m}$ & \cite{burgoyneemail} \\
Length of Gain Fibre & $L_g$ & $2$--$3 \text{ m}$ & \cite{burgoyne2014, peng, shtyrina, yarutkina} \\
Modulation Time & $T_M$ & $15$--$150 \text{ ps}$ & \cite{bohun, burgoyne2014, burgoyneemail} \\
Reflectivity of Optical Coupler & $R$ & $0.1$--$0.9$ & \cite{burgoyneemail, peng} \\
Saturation Energy & $\Es$ & $10^3$--$10^4 \text{ pJ}$ & \cite{burgoyneemail, usechak, yarutkina} \\
Small Signal Gain & $g_0$ & $1$--$10 \text{ m}^{-1}$ & \cite{burgoyneemail, yarutkina} \\
\hline
\end{tabular}
\caption{Orders of magnitude of various parameters.}
\label{tab:values}
\end{table}

\section{Combining the Effects}
\label{sec:effects}
In this model the pulse is iteratively passed through each process, the order of which must now be considered. We are most interested in the output of the laser cavity, and so we shall start with the loss component. Next the pulse is passed though the CFBG, as well as the modulator. Finally, the pulse travels through the gain fibre to be amplified, and then we consider the effect of the nonlinearity since this is the region where the power is maximized. Note that in general the functional operators of the components do not commute, and therefore the order is indeed important---in contrast to the previous models. This is especially the case of dispersion due to the Fourier transform. The pulse after one complete circuit of the laser cavity is then passed back in to restart the process. Functionally this can be denoted as
\begin{align*}
	\mathcal{L}(A) = F(G(M(D(L(A))))),
\end{align*}
where $\mathcal{L}$ is one loop of the laser. A solution to this model is one in which the envelope and chirp are unchanged after traversing every component in the cavity, that is, such that $\mathcal{L}(A) = A \textrm{e}^{i \phi}$---for some $\phi \in \mathbb{R}$.

\section{Solution to the Linear Model}
\label{sec:linear}
In the case of $b = 0$---which shall be referred to as the linear model---a solution can be found analytically. It is expected the solution will take the form of a Gaussian. There are a few reasons for this; the solution to the previous models were Gaussian \cite{cutler, siegman, kuizenga1970a, martinez1984, martinez1985} (see Section \ref{chap:meml}), the equilibrium shape will be highly correlated to the shape of the modulation function, and since a Gaussian is a fixed point of the Fourier transform \cite{gradshteyn} dispersion will not alter the envelope. \\

Consider the initial pulse
\begin{align*}
	A_0 = \sqrt{P} \exp \left( -(1 + iC) \frac{T^2}{2 \sigma^2} \right) \textrm{e}^{i \phi_0},
\end{align*}
where $P$ is the peak power, $C$ is the chirp, $\sigma^2$ is the variance, and $\phi_0$ is the initial phase. After passing through the optical coupler the pulse will simply decay to $A_1 = h A_0$. The pulse then enters the CFBG, where it will maintain its Gaussian shape, however, it will spread \cite{agrawal2013, ferreira, silfvast}. This can be written as
\begin{align*}
A_2 = \sqrt{P} h \zeta \exp \left( -(1 + i \widetilde{C}) \frac{T^2}{2 \widetilde{\sigma}^2} \right) \textrm{e}^{i(\phi_0 + \phi)},
\end{align*}
where $\widetilde{\sigma}^2$ denotes the resulting variance, $\widetilde{C}$ denotes the resulting chirp, and $\zeta$ is the reduction of the amplitude caused by the spread. Next, the pulse is modulated:
\begin{align*}
A_3 = \sqrt{P} h \zeta \exp \left( -(1 + i \widetilde{C}) \frac{T^2}{2 \widetilde{\sigma}^2} - \frac{T^2}{2} \right) \textrm{e}^{i(\phi_0 + \phi)}.
\end{align*}
Finally, the pulse travels through the gain fibre where it is amplified to
\begin{align*}
A_4 = \sqrt{P} \left( \frac{W(a E \textrm{e}^E)}{E} \right)^{1/2} h \zeta \exp \left( -(1 + i \widetilde{C}) \frac{T^2}{2 \widetilde{\sigma}^2} - \frac{T^2}{2} \right) \textrm{e}^{i(\phi_0 + \phi)},
\end{align*}
with $E$ the energy of the pulse as it enters the gain fibre. \\

In equilibrium it must be that $A_0 = A_4 \textrm{e}^{-i \phi}$. More explicitly, this gives three conditions:
\begin{subequations}
\begin{align}
\label{eq:energycond}
1 &= \left( \frac{W(a E \textrm{e}^E)}{E} \right)^{1/2} h \zeta, \\
\label{eq:varcond}
\frac{1}{\sigma^2} &= \frac{1}{\widetilde{\sigma}^2} + 1, \\
\label{eq:chirpcond}
\frac{C}{\sigma^2} &= \frac{\widetilde{C}}{\widetilde{\sigma}^2}.
\end{align}
\label{eq:linsys}
\end{subequations}

\subsection{Spread Due to Dispersion}
Each of these processes has a relatively straight forward effect, with the exception of dispersion.
\input{./Appendices/Dispersion}

>>>>

\chapter{A New Model}
\label{chap:linear}
In this chapter we shall derive our new model and solve it analytically in the linear case. To accomplish this we shall use the ideas of the previous functional models \cite{cutler, siegman, kuizenga1970a, kuizenga1970b, kuizenga1970, martinez1984, martinez1985, burgoyne2014}. To alleviate some of the pitfalls mentioned in the previous chapter, we shall include all five processes involved in the modification of the pulse within the laser cavity (gain, nonlinearity, loss, dispersion, and modulation). In addition to this, the functional operations associated with each component will be derived from \eqref{eq:nlse}, with the exception of the modulation in which we consider the exact functional form to be determined by the laser operator. \\

\section{Components}
\label{sec:comp}
We shall begin our analysis with the derivation of the functional operators for the five components. \\

\subsection{Gain}
\label{chap:gain}
Within the Er-doped gain fibre, the gain term is dominant, and equation~\eqref{eq:nlse} reduces to
\begin{align}
\label{eq:gainde}
\pdiff{A}{z} &= \frac{1}{2} g(A) A,
\end{align}
where $g(A)$ takes the form \cite{bohun, burgoyne2014, hausbook, haus1975, haus1991, haus1992, haus2000, kartner, peng, shtyrina, silfvast, usechak, yarutkina}
\begin{align}
\label{eq:energy}
	g(A) &= \frac{g_0}{1 + E / \Es},& E &= \int_{-\infty}^\infty |A|^2 \, \df T,
\end{align}
where $g_0$ is a small signal gain, $E$ is the energy of the pulse, and $\Es$ is the energy at which the gain begins to saturate.
Multiplying~\eqref{eq:gainde} by $\bar{A}$, the complex conjugate of $A$, yields
\begin{align*}
	\bar{A} \pdiff{A}{z} = \frac{1}{2}\frac{g_0 |A|^2}{1 + E / \Es},
\end{align*}
adding this to its complex conjugate gives
\begin{align}
\pdiff{|A|^2}{z} &= \frac{g_0 |A|^2}{1 + E / \Es}.
\label{eq:diffea}
\end{align}
After integrating this becomes
\begin{align}
\diff{E}{z} &= \frac{g_0 E}{1 + E / \Es}.
\label{eq:diffez}
\end{align}
For $E \ll \Es$ the energy will grow exponentially, whereas for $E \gg \Es$ the gain has saturated and so the growth is linear. To obtain a closed form solution, \eqref{eq:diffez} is integrated over a gain fibre of length $L_g$, the length of the gain fibre, and assuming the energy increases from $E$ to $E_{\textrm{out}}$, then
\begin{align*}
	g_0 L_g = \log\frac{E_{\textrm{out}}}{E} + \frac{E_{\textrm{out}}-E}{\Es},
\end{align*}
and by exponentiating, rearranging, and applying $W$, the Lambert $W$ function\footnote{See Appendix \ref{chap:lambertw}.},
\begin{align*}
W \left(\frac{E}{\Es} \textrm{e}^{E/\Es} \textrm{e}^{g_0 L_g}\right) &= W\left(\frac{E_{\textrm{out}}}{\Es} \textrm{e}^{E_{\textrm{out}}/\Es}\right) = \frac{E_{\textrm{out}}}{\Es},
\end{align*}
by \eqref{eq:lambertw}. This results in the closed form expression
\begin{align*}
E_{\textrm{out}} = \Es W \left( \frac{E}{\Es} \textrm{e}^{E/\Es} \textrm{e}^{g_0 L_g} \right).
\end{align*}
Notice from \eqref{eq:diffea} and \eqref{eq:diffez} that
\begin{align*}
\frac{1}{|A|^2}\pdiff{|A|^2}{z} &= \frac{1}{E}\diff{E}{z}
\end{align*}
and so, $E \sim |A|^2$, therefore the gain in terms of the amplitude is given by
\begin{align*}
G(A;E) &= \left(\frac{E_{\textrm{out}}}{E}\right)^{1/2}A = \left( \frac{\Es}{E} W \left( \frac{E}{\Es} \textrm{e}^{E/\Es} \textrm{e}^{g_0 L_g} \right) \right)^{1/2} A
\end{align*}
after passing through the gain fibre once.

\subsection{Fibre Nonlinearity}
The nonlinearity of the fibre arises from the parameter $\gamma$; in regions where this effect is dominant expression~\eqref{eq:nlse} becomes
\begin{align}
\label{eq:fibrediff}
	\pdiff{A}{z} - i \gamma |A|^2 A = 0.
\end{align}
This expression can be manipulated in a similar manner to the gain to show $\frac{\partial}{\partial z} |A|^2 = 0$, suggesting the \acrshort{wkb} ansatz $A(z,T) = A_0(T) e^{i \varphi(T,z)}$. Substituting this representation into~\eqref{eq:fibrediff} and setting $\varphi(T,0)=0$ gives $\varphi(T,z) = \gamma |A|^2 z$. For a fibre of length $L_f$ the effect of the nonlinearity is given by the map $A \mapsto F(A)$, where
\begin{align*}
F(A) &= A \textrm{e}^{i \gamma |A|^2 L_f}.
\end{align*}
This is also frequently called the Kerr nonlinearity or Kerr effect after John Kerr who discovered the effect in 1875. As we shall see in Section \ref{chap:energy}, this is what is responsible for \acrlong{spm} and ultimately the degradation of the pulse \cite{desurvire, dunlop, martinez1984, tamura1996}.

\subsection{Loss}
Two sources of loss exist within the laser circuit: the loss due to the output coupler and the optical loss due to absorption and scattering. The first case is simply a scalar multiplication depending on the reflectivity of the output coupler. The loss due to absorption and scattering can be derived from \eqref{eq:nlse},
\begin{align*}
\pdiff{A}{z} = - \alpha A.
\end{align*}
Combining these two effects gives a loss that takes the form
\begin{align*}
L(A) &= (1 - R) \textrm{e}^{- \alpha L_T}A,
\end{align*}
where $R$ is the reflectivity of the output coupler, and $L_T$ is the total length of the laser circuit.

\subsection{Dispersion}
Within the laser cavity, the dispersion is dominated by the \gls{cfbg}. In comparison, the dispersion due to the fibre is negligible. The dispersive terms of \eqref{eq:nlse} give
\begin{align}
\label{eq:dispde}
	\pdiff{A}{z} = -i \frac{\beta_2}{2} \pdiff[2]{A}{T},
\end{align}
and since dispersion acts in the frequency domain it is convenient to take the Fourier transform\footnote{We use $\displaystyle \FT{f} = (2 \pi)^{1/2} \int_\mathbb{R} f(x) \textrm{e}^{i \omega x} \df x$ as the definition of the Fourier transform.} of \eqref{eq:dispde} \cite{debnath, gradshteyn}, giving
\begin{align*}
	\pdiff{}{z}\FT{A} &= i\frac{\omega^2}{2}\beta_2 \FT{A}.
\end{align*}
The effect of dispersion is then by the map
\begin{align*}
D(A) &= \FTi{\textrm{e}^{i \omega^2 L_D\beta_2/2} \FT{A}},
\end{align*}
where $L_D$ is the length of the dispersive medium. \\

\subsection{Modulation}
In \eqref{eq:meml}, the amount of modulation is characterized by the parameter $M_s$ through the term $\frac{1}{2} M_s T^2 A$. In this new model, the modulation is considered to be applied externally through its action on the spectrum. For simplicity the representation is taken as the Gaussian
\begin{align*}
M(A) &= \textrm{e}^{-T^2 / 2 T_M^2} A,
\end{align*}
where $T_M$ is a characteristic width of the modulation. \\

Despite assuming a Gaussian modulation, the solution presented in Section \ref{sec:linear} can be generalized to any modulation function in $L^2(\mathbb{R})$ since Gaussians in fact span $L^2(\mathbb{R})$ as shown in Appendix \ref{chap:gauss}. \\

%\input{./Appendices/Hermite}

@@

The structure of each process of the laser can be better understood by re-scaling the time, energy, and amplitude. Nominal values for tuneable lasers are shown in Table~\ref{tab:values}. Knowing experimental durations and energies, the table suggests the convenient scalings:
\begin{align}
	T &= T_M \widetilde{T},& E &= \Es \widetilde{E},& A &= \left( \frac{\Es}{T_M} \right)^{1/2} \widetilde{A}.
\end{align}
Revisiting each process map shows each process has a characteristic non-dimensional parameter. The new mappings, after dropping the tildes, are
\begin{subequations}
\label{eq:effects}
\begin{align}
\label{eq:gain}
G(A) &= \left(E^{-1} W \left( a E \textrm{e}^{E}\right) \right)^{1/2} A, \\
\label{eq:fibre}
F(A) &= A \textrm{e}^{i b |A|^2}, \\
\label{eq:loss}
L(A) &= h A, \\
\label{eq:disp}
D(A) &= \FTi{\textrm{e}^{i s^2 \omega^2} \FT{A}}, \\
\label{eq:mod}
M(A) &= \textrm{e}^{-T^2 / 2} A,
\end{align}
\end{subequations}
with the four dimensionless parameters, as defined by the values in Table \ref{tab:values},
\begin{equation}
\begin{aligned}
	a &= \textrm{e}^{g_0 L_g} \sim 8 \times 10^3,& \qquad h &= (1 - R) \textrm{e}^{-\alpha L} \sim 0.04\\
	b &= \gamma L_f \frac{\Es}{T_M} \sim 1,& \qquad s &= \sqrt{\frac{\beta_2 L_D}{2 T_M^2}} \sim 0.2,
\label{eq:ndparam}
\end{aligned}
\end{equation}
which characterize the behaviour of the laser. Notice that the modulation is characterized by $T_M$, and each other process has it's own independent non-dimensional parameter.

\begin{table}[tbp]
\centering
\begin{tabular}{lcll}
Parameter & Symbol & Value & Sources \\
\noalign{\global\arrayrulewidth=1.5pt}\hline
Absorption of Fibre\footnotemark & $\alpha$ & $10^{-4}$--$0.3\text{ m}^{-1}$  & \cite{burgoyneemail, shtyrina, usechak, tomlinson, yarutkina} \\
Fibre Dispersion & $\beta_2^f$ & $-50$--$50 \text{ ps}^2/ \text{km}$ & \cite{agrawal2002, agrawal2013, burgoyne2014, litchinitser, peng, yarutkina} \\
Fibre Nonlinearity & $\gamma$ & $0.001$--$0.01 \text{ W}^{-1} \text{m}^{-1}$ & \cite{agrawal2013, finot, usechak, yarutkina} \\
Grating Dispersion & $\beta_2^g L_D$ & $10$--$2000 \text{ ps}^2$ & \cite{agrawal2002, agrawal2013, burgoyne2014, li} \\
Length of Cavity & $L_T$ & $10$--$100 \text{ m}$ & \cite{burgoyneemail, peng, tamura1996} \\
Length of Fibre & $L_f$ & $0.15$--$1 \text{ m}$ & \cite{burgoyneemail} \\
Length of Gain Fibre & $L_g$ & $2$--$3 \text{ m}$ & \cite{burgoyne2014, peng, shtyrina, tamura1993, yarutkina} \\
Modulation Time & $T_M$ & $15$--$150 \text{ ps}$ & \cite{bohun, burgoyne2014, burgoyneemail} \\
Reflectivity of Optical Coupler & $R$ & $0.1$--$0.9$ & \cite{burgoyneemail, peng, li, tamura1993, tamura1996, yamashita} \\
Saturation Energy & $\Es$ & $10^3$--$10^4 \text{ pJ}$ & \cite{burgoyneemail, usechak, yarutkina} \\
Small Signal Gain & $g_0$ & $1$--$10 \text{ m}^{-1}$ & \cite{burgoyneemail, yarutkina} \\
\hline
\end{tabular}
\caption{Range of variation of various parameters.}
\label{tab:values}
\end{table}

\footnotetext{Fibre loss is typically reported as $\sim0.5$ dB/km.}

\section{Combining the Effects of Each Block of the Model}
\label{sec:effects}
In this model the pulse is iteratively passed through each process, the order of which must now be considered. We are most interested in the output of the laser cavity, and so we shall start with the loss component. Next the pulse is passed though the \gls{cfbg}, as well as the modulator. Finally, the pulse travels through the gain fibre to be amplified, and then we consider the effect of the nonlinearity since this is the region where the power is maximized. Note that in general the functional operators of the components do not commute, and therefore the order of the components is indeed important---in contrast to the previous models. This is especially the case of dispersion as realized through the Fourier transform. The pulse after one complete circuit of the laser cavity is then passed back in to restart the process. Functionally this can be denoted as
\begin{align}
	\mathcal{L}(A) = F(G(M(D(L(A))))),
\end{align}
where $\mathcal{L}$ is one loop of the laser. A steady solution to this model is one in which the envelope and chirp are unchanged after traversing every component in the cavity, that is, such that $\mathcal{L}(A) = A \textrm{e}^{i \phi}$---for some $\phi \in \mathbb{R}$. The phase shift is allowable because it is immeasurable in the laboratory---only the envelope, chirp, and variance can be measured.

\section{Solution to the Linear Model}
\label{sec:linear}
In reality the nonlinearity parameter, $b$, is of order unity by \eqref{eq:ndparam}, but we shall start our investigation with the simpler case of $b = 0$---which shall be referred to as the linear model. In this case, a solution can be found analytically. It is expected the solution will take the form of a Gaussian. There are a few reasons for this; the solution to the previous models were Gaussian \cite{cutler, siegman, kuizenga1970a, martinez1984, martinez1985} (see Section \ref{chap:meml}), the equilibrium shape will be highly correlated to the shape of the modulation function, and since a Gaussian is a fixed point of the Fourier transform \cite{gradshteyn}, dispersion will not alter the envelope. \\

\begin{figure}[tbp]
\centering
\input{./Figures/Sample_Gauss}
\caption{Sample realization of \eqref{eq:A0}.}
\label{fig:samplegauss}
\end{figure}

To compute $\mathcal{L}(A)$, consider the initial pulse
\begin{align}
	A_0 = \sqrt{P} \exp \left( -(1 + iC) \frac{T^2}{2 \sigma^2} \right) \textrm{e}^{i \phi_0},
\label{eq:A0}
\end{align}
where $P$ is the peak power, $C$ is the chirp\footnote{As discussed in Section \ref{sec:fbg}.}, $\sigma^2$ is the variance, and $\phi_0$ is the initial phase. An example of such a pulse is shown in Figure \ref{fig:samplegauss}. This form is chosen because it matches the envelope and linear chirping found by experiment. After passing through the optical coupler the pulse will simply decay to $A_1 = L(A_0) = h A_0$. The pulse then enters the \gls{cfbg}, where it will maintain its Gaussian shape, however, it will spread \cite{agrawal2013, ferreira, silfvast}. This can be written as
\begin{align}
A_2 = D(A_1) = \sqrt{P} h \zeta \exp \left( -(1 + i \widetilde{C}) \frac{T^2}{2 \widetilde{\sigma}^2} \right) \textrm{e}^{i(\phi_0 + \phi)},
\end{align}
where $\widetilde{\sigma}^2$ denotes the resulting variance, $\widetilde{C}$ denotes the resulting chirp, $\zeta$ is the reduction of the amplitude caused by the spread, and $\phi$ is the inherited phase shift from dispersion. At this point $\widetilde{\sigma}^2$, $\widetilde{C}$, $\zeta$, and $\phi$ remain undetermined, and are used as placeholders for convenience. Next, the pulse is modulated:
\begin{align}
A_3 = M(A_2) = \sqrt{P} h \zeta \exp \left( -(1 + i \widetilde{C}) \frac{T^2}{2 \widetilde{\sigma}^2} - \frac{T^2}{2} \right) \textrm{e}^{i(\phi_0 + \phi)}.
\end{align}
Finally, the pulse travels through the gain fibre where it is amplified to
\begin{align}
A_4 = G(A_3) = \mathcal{L}(A_0) = \sqrt{P} \left( \frac{W(a E \textrm{e}^E)}{E} \right)^{1/2} h \zeta \exp \left( -(1 + i \widetilde{C}) \frac{T^2}{2 \widetilde{\sigma}^2} - \frac{T^2}{2} \right) \textrm{e}^{i(\phi_0 + \phi)},
\label{eq:A4}
\end{align}
with $E$ the energy of the pulse as it enters the gain fibre. \\

In equilibrium it must be that $A_0 = A_4 \textrm{e}^{-i \phi} = \mathcal{L}(A_0) \textrm{e}^{- i \phi}$ so that after a single loop of the cavity the variance, chirp, and envelope remain unchanged. More explicitly, this gives three conditions:
\begin{subequations}
\begin{align}
\label{eq:energycond}
1 &= \left( \frac{W(a E \textrm{e}^E)}{E} \right)^{1/2} h \zeta, \\
\label{eq:varcond}
\frac{1}{\sigma^2} &= \frac{1}{\widetilde{\sigma}^2} + 1, \\
\label{eq:chirpcond}
\frac{C}{\sigma^2} &= \frac{\widetilde{C}}{\widetilde{\sigma}^2}.
\end{align}
\label{eq:linsys}
\end{subequations}

\subsection{Spread of the Pulse Due to Dispersion}
Each of these processes has a relatively straight forward effect, with the exception of dispersion.
\input{./Appendices/Dispersion}

\subsection{Equilibrium Shape of the Pulse}
Now that the effect of dispersion is know analytically, the system \eqref{eq:linsys} can be solved. The out-going variance and chirp can be eliminated from this system of equations using \eqref{eq:varcond} and \eqref{eq:chirpcond}:
\begin{align*}
\widetilde{\sigma}^2 &= \frac{\sigma^2}{1 - \sigma^2}, & \widetilde{C} &= C \frac{1}{1 - \sigma^2}.
\end{align*}
Combining this first expression with \eqref{eq:varout} and expanding the square, we have
\begin{align*}
\frac{\sigma^4}{1 - \sigma^2} = \sigma^4 + 4 C^2 s^4 + 4 C s^2 \sigma^2 + 4s^4,
\end{align*}
or written as a polynomial in $\sigma$,
\begin{align}
0 = \sigma^6 + 4 C s^2 \sigma^4 + \left( 4s^4 (C^2 + 1) - 4 C s^2 \right) \sigma^2 - 4s^4 (C^2 + 1).
\label{eq:midvar}
\end{align}
The chirp can now be eliminated using \eqref{eq:chirpout}, the $1+C^2$ can be reduced in order by noticing that
\begin{align*}
1 + C^2 &= \frac{\sigma^4}{2s^2(1 - \sigma^2)} C.
\end{align*}
Solving for $C$ gives
\begin{align}
\label{eq:equilchirp}
C &= \frac{\sigma^4}{2s^2(1 - \sigma^2)} \pm \sqrt{\frac{\sigma^8}{16s^4(1 - \sigma^2)^2} - 1} = \frac{\sigma^4 \pm \sqrt{\sigma^8 - 16s^4(1 - \sigma^2)^2}}{4s^2(1 - \sigma^2)},
\end{align}
and after simplifying \eqref{eq:midvar} algebraically, we arrive at
\begin{align}
\frac{\sigma^6}{2 - \sigma^2} &= \mp \sqrt{\sigma^8 - 16s^4(1 - \sigma^2)^2}.
\label{eq:varroot}
\end{align}
As we shall see, $\sigma$ is strictly less than $1$ at equilibrium, and so, notice that the left hand side of this expression is strictly positive. Therefore, only the negative root of \eqref{eq:equilchirp} will yield a solution. Moreover, Martinez, Fork, and Gordon \cite{martinez1984} found two solutions to their discrete model (Section \ref{sec:discrete}) analytically, but showed that one was unstable---consistent with experiments. After squaring each side of \eqref{eq:varroot} we obtain the biquartic equation
\begin{align}
\label{eq:var}
\left( \sigma^2 \right)^4 + 4 s^4 \left( \sigma^2 \right)^3 - 20 s^4 \left( \sigma^2 \right)^2 + 32 s^4 \left( \sigma^2 \right) - 16 s^4 = 0.
\end{align}
Since this is a quartic in $\sigma^2$ this can be solved analytically; the (positive, real) solution is
\begin{align}
\label{eq:equilvar}
\sigma^2 &= \sqrt{2} s \left( s^6 + 3s^2 + \sqrt{4 + s^4} \left( 1 + s^4 \right) \right)^{1/2} - s^4 - s^2 \sqrt{4 + s^4}.
\end{align}
A sample solution for the linear model is shown in Figure \ref{fig:linear}. To reiterate, the envelope at equilibrium is Gaussian, and therefore, so too is the Fourier transform. Also, the pulse is linearly chirped. \\
